---
title: Blog Post 7
author: Transformers
date: '2022-12-02'
slug: []
categories: []
tags: []
description: ~
toc: yes
authors: []
series: []
lastmod: '2022-12-02T07:38:30-04:00'
featuredVideo: ~
featuredImage: ~
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(modelr))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(tmap))
suppressPackageStartupMessages(library(dplyr))
load(here::here("dataset", "abortionDataClean.RData"))
load(here::here("dataset", "MaxStateAff.RData"))
us_states <- 
  st_read(here::here("dataset","cb_2019_us_state_20m/cb_2019_us_state_20m.shp"))
geog_loc <- 
  read_csv(here::here("dataset","StateGeography.csv"))
us_regions <- 
  st_read(here::here("dataset","cb_2018_us_division_20m/cb_2018_us_division_20m.shp"))
```

# Ideas for Continued Exploratory Data Analysis:

## Average Pregnancy Rate By State Affiliation

```{r, echo = FALSE, warning = FALSE, message = FALSE}
abortionDataClean %>% 
  group_by(state) %>% 
  summarise(AveragePregnancyRate = mean(PregnancyRate, na.rm=TRUE)) %>% 
  inner_join(MaxStateAff) %>% 
  ggplot() + 
  stat_summary(aes(x = reorder(state, AveragePregnancyRate), y = AveragePregnancyRate, fill = as.factor(Affiliation)), geom = "bar") + 
  coord_flip() + 
  scale_fill_viridis_d(option = "cividis") + 
  labs(title = 'Average Pregnancy Rate by State Affiliation', x = 'State\n', y = '\nAverage Pregnancy Rate', fill = "Affliation") + 
  ggthemes::theme_economist() +
  theme(axis.text.y = element_text(size = 7), axis.title = element_text(size = 10), legend.position = 'right', legend.text = element_text(size = 8))
```

Stuff about Average pregnancy Rate by State Affiliation

## Average Abortion Rate by Age group 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
abortionDataClean %>% 
  group_by(year, Age) %>% 
  summarise(AverageAbortionRate = mean(AbortionRate, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = AverageAbortionRate, color = as.factor(Age))) + 
  geom_line() + 
  labs(title = 'Average Abortion Rate by Age Group from 1973 to 2017', x = 'Year', y = 'Average Abortion Rate', color = "Age Group") + 
  ggthemes::theme_economist() +
  theme(legend.position = 'right', legend.text = element_text(size = 8)) +
  scale_color_viridis_d(option = "turbo")
```

Stuff about Average Abortion Rate by Age group

## Average Pregnancy Rate by Age group for Different Political Affiliations

```{r, echo = FALSE, warning = FALSE, message = FALSE}
abortionDataClean %>% 
  group_by(state, year, Age) %>% 
  summarise(AveragePregRate = mean(PregnancyRate, na.rm = TRUE)) %>% 
  ungroup() %>% 
  inner_join(MaxStateAff) %>% 
  filter(Affiliation %in% c("DEMOCRAT", "REPUBLICAN")) %>% 
  ggplot(aes(x = year, y = AveragePregRate, color = as.factor(Age))) + 
  geom_line() + 
  facet_wrap(~ as.factor(Affiliation)) +
  labs(title = 'Average Pregnancy Rate by Age Group for Different Political Affiliations', x = 'Year', y = 'Average Pregnancy Rate', color = "Age Group")
```

Stuff about Average Pregnancy Rate by Age Group for Differnet Political Affiliations

## Average Abortion Rate by Age group for Different Political Affiliations

```{r, echo = FALSE, warning = FALSE, message = FALSE}
abortionDataClean %>% 
  group_by(state, year, Age) %>% 
  summarise(AverageAbortionRate = mean(AbortionRate, na.rm = TRUE)) %>% 
  ungroup() %>% 
  inner_join(MaxStateAff) %>% 
  filter(Affiliation %in% c("DEMOCRAT", "REPUBLICAN")) %>% 
  ggplot(aes(x = year, y = AverageAbortionRate, color = as.factor(Age))) + 
  geom_line() + 
  facet_wrap(~ as.factor(Affiliation)) +
  labs(title = 'Average Abortion Rate by Age Group for Different Political Affiliations', x = 'Year', y = 'Average Abortion Rate', color = "Age Group")
```

Stuff about Average Abortion Rate by Age Group for Different Political Affiliations

## Political Affiliation by State
```{r readshapefile, echo = FALSE, warning = FALSE, message = FALSE}
not_included <-
  c("Guam", "Commonwealth of the Northern Mariana Islands",
    "American Samoa", "Puerto Rico", "United States Virgin Islands")

us_states <- 
  us_states %>%
  filter(!(NAME %in% not_included))
```

```{r convert to EPSG 2163, echo = FALSE}
epsg_us <- 
  2163

us_states <- 
  st_transform(us_states, epsg_us)
```

```{r join US states, geographic location and state affiliation datasets, warning=FALSE, message=FALSE, echo = FALSE,}
states_geog <- 
  inner_join(us_states, geog_loc, by = c("STUSPS" = "State", "NAME" = "FullStateName"))

states_geog_aff <- 
  inner_join(states_geog, MaxStateAff, by = c("STUSPS" = "state"))
```

```{r calculate overall abortion rate for each state, echo = FALSE,}
state_abortion_rate <- 
  abortionDataClean %>% 
  replace(is.na(.), 0) %>% 
  group_by(state) %>% 
  select(state, Abortions, Population) %>% 
  mutate(state, total_population = sum(Population), total_abortions = sum(Abortions), overall_abortion_rate = total_abortions/total_population) %>% 
  select(state, total_abortions, total_population, overall_abortion_rate)

state_abortion_rate <- 
  state_abortion_rate[!duplicated(state_abortion_rate$state), ]
```

```{r join overall abortion rate dataset with state, geographic location and political affiliation dataset, warning=FALSE, message=FALSE, echo = FALSE,}
state_data <- 
  inner_join(states_geog_aff, state_abortion_rate, by = c("STUSPS" = "state")) %>%
  select(NAME, STUSPS, GeographicRegion, Affiliation, overall_abortion_rate, geometry)
```

```{r plot map by state, fig.width=80, fig.height=80, warning=FALSE, message=FALSE, echo = FALSE,}
tmap_mode("view")
colours <- 
  c("#00AEF3", "#808080", "#E81B23")

tm_shape(state_data) + 
  tm_polygons(col = "overall_abortion_rate", title = "Overall Abortion Rate by State") +
  tm_layout(frame = FALSE, legend.outside = TRUE, legend.title.size = 2, legend.text.size = 1.5)
  
tm_shape(state_data) + 
  tm_polygons(col = "Affiliation", title = "Political Affiliation by State", palette = colours) +
  tm_layout(frame = FALSE, legend.outside = TRUE, legend.title.size = 2, legend.text.size = 1.5)
```

Stuff about map 1

## Overall Abortion Rate by Region

```{r load region data, warning=FALSE, message=FALSE, echo = FALSE}
us_regions <- 
  us_regions %>% 
  st_transform(epsg_us)
```

```{r get overall abortion data by region, warning=FALSE, message=FALSE, echo = FALSE}
region_abortion_rate <- 
  inner_join(states_geog_aff, abortionDataClean, by = c("STUSPS" = "state")) %>% 
  replace(is.na(.), 0) %>% 
  group_by(GeographicRegion) %>% 
  select(GeographicRegion, Abortions, Population) %>% 
  mutate(GeographicRegion, total_population = sum(Population), total_abortions = sum(Abortions), overall_abortion_rate = total_abortions/total_population) %>% 
  select(GeographicRegion, total_abortions, total_population, overall_abortion_rate) %>% 
  st_drop_geometry()

region_abortion_rate <- 
  region_abortion_rate[!duplicated(region_abortion_rate$GeographicRegion), ]

region_abortion_rate <- 
  region_abortion_rate[!duplicated(region_abortion_rate$GeographicRegion), ]

region_abortion_rate <- 
  st_set_geometry(region_abortion_rate, us_regions$geometry)
```

```{r plot map by geographic location, fig.width=80, fig.height=80, warning=FALSE, message=FALSE, echo = FALSE}
region_data <- 
  region_abortion_rate %>% 
   select(GeographicRegion, overall_abortion_rate, geometry)

tmap_mode("view")

tm_shape(region_data) + 
  tm_polygons(col = "overall_abortion_rate", title = "Overall Abortion Rate by Region") +
  tm_shape(state_data) + 
  tm_borders() +
  tm_layout(frame = FALSE, legend.outside = TRUE, legend.title.size = 2, legend.text.size = 1.5)
```

Stuff about map 2