---
title: Blog Post 7
author: Transformers
date: '2022-12-02'
slug: []
categories: []
tags: []
description: ~
toc: yes
authors: []
series: []
lastmod: '2022-12-02T07:38:30-04:00'
featuredVideo: ~
featuredImage: ~
---

# Ideas for Continued Exploratory Data Analysis:
*This post is to discuss how we are continuing our exploratory data analysis.


```{r setup, warning=FALSE, message=FALSE}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(tmap))
suppressPackageStartupMessages(library(dplyr))
```

```{r read shape file}

us_states <- st_read("../../../dataset/cb_2019_us_state_20m/cb_2019_us_state_20m.shp", quiet = TRUE)

not_included <-
  c("Guam", "Commonwealth of the Northern Mariana Islands",
    "American Samoa", "Puerto Rico", "United States Virgin Islands")

us_states <- us_states %>%
  filter(!(NAME %in% not_included))

print(us_states)
```

```{r convert to EPSG 2163}

epsg_us <- 2163
us_states <- st_transform(us_states, epsg_us)
```

```{r join geographic location with states data, warning=FALSE, message=FALSE}

geog_loc <- read_csv("../../../dataset/StateGeography.csv")

states_geog <- inner_join(us_states, geog_loc, by = c("STUSPS" = "State", "NAME" = "FullStateName"))

print(states_geog)
```

```{r plot map, fig.width=80, fig.height=80}

tm_shape(states_geog) + tm_polygons() + 
  tm_text("STUSPS") +
  tm_layout(frame = FALSE, scale = 5)
```

# Polishing our Visualizations and Tables: 
* Because presentation is important when communicating science, we know that it is crucial to spend time adding titles, captions, and other tools make our graphs as clear and user-friendly as possible. 

*Think about titles, captions, annotations, highlights. Look into ggpubr for tools to improve your figures and look into some of the options for displaying tables using https://gallery.htmlwidgets.org/.
